Offline Mode for Corporate Mobile Client Security Architecture

### 2016-10-23 docs/papers/2016MONE/2016MONE.tex
several updates

\section{The mobile security architecture}
\label{sec_the_client_security_architecture}
The approach proposed in this work describes and implements the complete lifecycle of the mobile client security infrastructure. The mobile security processes depend on the key expiry period, and are used to access the protected storage. Once the user keys expire, the user is requested to enter his valid credentials, i.e. PIN and password. The mobile client then sends the credentials to the server for verification. Once the new set of access keys is received, the user can access the protected files in the offline mode, without the access to the server. This means that no further communication with the server is needed until the key expires. The core set of functions and protocols can be divided into three sets of operations as shown in Figure \ref{fig:1}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=5cm]{figures/coresetoffunctionsandprotocols.eps}
	\caption{The core set of functions and protocols of the mobile cloud security infrastructure}
	\label{fig:1}
\end{figure}

These operations function based on the mobile client architecture, which is depicted in Figure \ref{fig:2}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=8cm]{figures/mobileclientarchitecture.eps}
	\caption{The Mobile Client Architecture}
	\label{fig:2}
\end{figure}

@@

The cryptographic functions include the decryption (for AES and ABE) key restoration for SSS and EKE token generation procedure. While the AES-ABE hybrid encryption scheme functions both in the online and offline modes, the SSS is used only in the offline mode, while the EKE generation and executions is used in the online mode. These operations are described in Subsections \ref{sec_aes}-\ref{sec_secret} and are part of the Figure \ref{fig:1}.

@@

The proposed architecture for offline mobile security is represented in Figure \ref{fig:5}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=8cm]{figures/offlinesecurityarchitecture.eps}
	\caption{Proposed Architecture for Offline Mobile Security }
	\label{fig:5}
\end{figure}

The infrastructure in Figure \ref{fig:5} includes:

@@

The mobile client performs all the cryptographic calculations in the shadow. These calculations include the key storage, key restoring, decryption and showing the decrypted files in the mobile client area. Note that the mobile client does not check the password or PIN validity as it does not hold the verification proof for the above. Also, the mobile client keeps track of the user activities and the key expiry. The complete list of the mobile client activities is presented in the Figure \ref{fig:6}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=8cm]{figures/offlinemodeoperations.eps}
	\caption{Offline Mode Operations}
	\label{fig:6}
\end{figure}

To finalize the description of the offline mode, Figure \ref{fig:7} shows the complete workflow of the proposed mobile application in the offline-mode.

\begin{figure}[h!]
	\centering
	\includegraphics[width=8cm]{figures/offlineauthorizationworkflow.eps}
	\caption{Offline mode authorization workflow}
	\label{fig:7}
\end{figure}

@@

In the proposed client security architecture, the Threat Intelligence Manager (TIM) is responsible for receiving logged user operations, perform feature extraction, data modeling and malicious behavior analysis in order to identify possible threats, in offline mode. Figure \ref{fig:8} depicts the TIM for offline behavioral analysis.

\begin{figure}[h!]
	\centering
	\includegraphics[width=8cm]{figures/mosworkflow.eps}
	\caption{The Threat Intelligence Manager Workflow}
	\label{fig:8}
\end{figure}

As depicted in Figure \ref{fig:8}, users request operations are logged so that the main features can be monitored in the mobile client. The user behavior, trying or effectively executing operations, shall be incrementally captured and logged, making possible to monitor the main features that can reveal malicious behaviors, as well as to identify unexpected behaviors that can reveal possible threats. Therefore, the user operations are monitored by the client app, which sends the information to the User Activity Logging (UAL).

@@

\section{The algorithms, key usage and data protection methods}
\label{sec_algorithms}
In the proposed approach, the kernel encryption scheme in the mobile client is a combination of several methods of security. The files are encrypted with 128/256-bit AES, while the permanent file keys are encrypted with the attribute-based encryption. The set of expiring ABE keys corresponding to the set of files accessible by user in encrypted with a single expiring AES key (KEY\_SET\_KEY). This key is expiring and is splited by the server into 4 parts (2 are stored on the device and 2 belong to the user) by the method of secure secret sharing. The encryption workflow is outlined in the Figure \ref{fig:9}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=8cm]{figures/encryptionworkflow.eps}
	\caption{Encryption workflow}
	\label{fig:9}
\end{figure}

The key hierarchy is explained by Figure \ref{fig:10} where the key represented in Figure \ref{fig:9} serves for decrypting the key for each level represented in Figure \ref{fig:10}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=8cm]{figures/encryptionkeyhierarchy.eps}
	\caption{Encryption key hierarchy}
	\label{fig:10}
\end{figure}

@@

The scheme of the server-side AES encryption is represented in Figure \ref{fig:11}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=8cm]{figures/server-sideencryption.eps}
	\caption{Server-side encryption}
	\label{fig:11}
\end{figure}

>>>>

\section{The mobile security architecture}
\label{sec_the_client_security_architecture}
The approach proposed in this work describes and implements the complete lifecycle of the mobile client security infrastructure. The mobile security processes depend on the key expiry period, and are used to access the protected storage. Once the user keys expire, the user is requested to enter his valid credentials, i.e. PIN and password. The mobile client then sends the credentials to the server for verification. Once the new set of access keys is received, the user can access the protected files in the offline mode, without the access to the server. This means that no further communication with the server is needed until the key expires. The core set of functions and protocols can be divided into three sets of operations as shown in Figure \ref{fig:01}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=5cm]{fig01.eps}
	\caption{The core set of functions and protocols of the mobile cloud security infrastructure}
	\label{fig:01}
\end{figure}

These operations function based on the mobile client architecture, which is depicted in Figure \ref{fig:02}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=8cm]{fig02.eps}
	\caption{The Mobile Client Architecture}
	\label{fig:02}
\end{figure}

@@

The cryptographic functions include the decryption (for AES and ABE) key restoration for SSS and EKE token generation procedure. While the AES-ABE hybrid encryption scheme functions both in the online and offline modes, the SSS is used only in the offline mode, while the EKE generation and executions is used in the online mode. These operations are described in Subsections \ref{sec_aes}-\ref{sec_secret} and are part of the Figure \ref{fig:01}.

@@

The proposed architecture for offline mobile security is represented in Figure \ref{fig:03}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=8cm]{fig03.eps}
	\caption{Proposed Architecture for Offline Mobile Security }
	\label{fig:03}
\end{figure}

The infrastructure in Figure \ref{fig:03} includes:

@@

The mobile client performs all the cryptographic calculations in the shadow. These calculations include the key storage, key restoring, decryption and showing the decrypted files in the mobile client area. Note that the mobile client does not check the password or PIN validity as it does not hold the verification proof for the above. Also, the mobile client keeps track of the user activities and the key expiry. The complete list of the mobile client activities is presented in the Figure \ref{fig:04}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=8cm]{fig04.eps}
	\caption{Offline Mode Operations}
	\label{fig:04}
\end{figure}

To finalize the description of the offline mode, Figure \ref{fig:05} shows the complete workflow of the proposed mobile application in the offline-mode.

\begin{figure}[h!]
	\centering
	\includegraphics[width=8cm]{fig05.eps}
	\caption{Offline mode authorization workflow}
	\label{fig:05}
\end{figure}

@@

In the proposed client security architecture, the Threat Intelligence Manager (TIM) is responsible for receiving logged user operations, perform feature extraction, data modeling and malicious behavior analysis in order to identify possible threats, in offline mode. Figure \ref{fig:06} depicts the TIM for offline behavioral analysis.

\begin{figure}[h!]
	\centering
	\includegraphics[width=8cm]{fig06.eps}
	\caption{The Threat Intelligence Manager Workflow}
	\label{fig:06}
\end{figure}

As depicted in Figure \ref{fig:06}, users request operations are logged so that the main features can be monitored in the mobile client. The user behavior, trying or effectively executing operations, shall be incrementally captured and logged, making possible to monitor the main features that can reveal malicious behaviors, as well as to identify unexpected behaviors that can reveal possible threats. Therefore, the user operations are monitored by the client app, which sends the information to the User Activity Logging (UAL).

@@

\section{The algorithms, key usage and data protection methods}
\label{sec_algorithms}
In the proposed approach, the kernel encryption scheme in the mobile client is a combination of several methods of security. The files are encrypted with 128/256-bit AES, while the permanent file keys are encrypted with the attribute-based encryption. The set of expiring ABE keys corresponding to the set of files accessible by user in encrypted with a single expiring AES key (KEY\_SET\_KEY). This key is expiring and is splited by the server into 4 parts (2 are stored on the device and 2 belong to the user) by the method of secure secret sharing. The encryption workflow is outlined in the Figure \ref{fig:07}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=8cm]{fig07.eps}
	\caption{Encryption workflow}
	\label{fig:07}
\end{figure}

The key hierarchy is explained by Figure \ref{fig:08} where the key represented in Figure \ref{fig:07} serves for decrypting the key for each level represented in Figure \ref{fig:08}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=8cm]{fig08.eps}
	\caption{Encryption key hierarchy}
	\label{fig:08}
\end{figure}

@@

The scheme of the server-side AES encryption is represented in Figure \ref{fig:09}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=8cm]{fig09.eps}
	\caption{Server-side encryption}
	\label{fig:09}
\end{figure}
